---
title: "2019 March Madness Bracket"
output: html_notebook
---
```{r}
pacman::p_load(tidyverse, here, mRchmadness)
```

Since I don't have time to run my own model, I'll start with 538's predictions.

```{r}
if(!file.exists("fivethirtyeight_ncaa_forecasts.csv")) {
    download.file("https://projects.fivethirtyeight.com/march-madness-api/2019/fivethirtyeight_ncaa_forecasts.csv",
                  "fivethirtyeight_ncaa_forecasts.csv")
}
```

```{r}
fte_predictions <- read_csv("fivethirtyeight_ncaa_forecasts.csv") %>%
    filter(gender == "mens")
fte_predictions 
```

Setup bracket pool point system.

```{r}
points <- tibble(round = grep("^rd", names(fte_predictions), value = T),
       base_points = c(0,1,2,4,8,16,32))
```

```{r}
expected_points <- fte_predictions %>%
    gather("round", "odds", starts_with("rd")) %>%
    left_join(points) %>%
    mutate(win_points = base_points + as.numeric(team_seed),
           expected_points = win_points * odds)

expected_points[expected_points$round == "rd1_win", c("win_points", "expected_points")] <- 0
```

```{r}
expected_points %>%
    mutate(team_name = fct_reorder(team_name, expected_points, sum)) %>%
    ggplot(aes(x = team_name, y = expected_points, fill = round)) +
    geom_col() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1))
```

```{r}
library("mRchmadness")
```

```{r}
head(pred.538.men.2018)
```

Who should we guess at for play in games?  Let's make a list of the losers to filter out

```{r}
play_in_losers <- fte_predictions %>%
    filter(grepl("a|b", team_seed)) %>%
    mutate(team_seed = str_replace(team_seed, "a|b", "")) %>%
    group_by(team_region, team_seed) %>%
    arrange(team_region, team_seed, rd1_win) %>%
    summarise_all(first) %>%
    ungroup() %>%
    select(team_id, team_name)
```

Now let's make a 2019 538 prediction data set.

```{r}
pred_538_men_2019 <- fte_predictions %>%
    anti_join(play_in_losers) %>%
    mutate(team_name = fct_inorder(team_name)) %>%
    select(name = team_name, 
           round1 = rd2_win,
           round2 = rd3_win,
           round3 = rd4_win,
           round4 = rd5_win,
           round5 = rd6_win,
           round6 = rd7_win)
```

Recreate simulation function...

```{r}
set.seed(1)

rownames(pred_538_men_2019) <- pred_538_men_2019$name


```

Create derived probabilites

```{r}

prob_derived <- pred_538_men_2019 %>%
    #select just the round columns
    select(starts_with("round")) %>%
    #make all values negative
    mutate_all(funs(.*-1)) %>%
    #create a matrix of differences
    apply(MARGIN = 1, FUN = diff) %>%
    #transpose the matrix
    t() %>%
    cbind(pred_538_men_2019$round6)
    
names(prob_derived) <- paste0("round", 1:6)

```

Set up the variable to stoare the path (in terms of game #) that each eam would need ot take to et to tthe championship. Ensures winner of round r won r-1

```{r}
path <- matrix(0, 64, 6)
rownames(path) <- pred_538_men_2019$name
```

Simulate the winner of each round

```{r}
outcome <- matrix('', 63, 1)

for (r in 1:6) {
    (path[,r] <- max(path) + rep(seq(2^(6-r)), each = 2^r))
    (groups <- split(pred_538_men_2019$name, f = path[,r]))
    (outcome[round == r, ] <- t(sapply(groups, function(i) {
        sample(i, 1, prob = prob_derived[i,r], replace = TRUE)})))
}

path
groups
```


for (r in 2:6) {
    #Identify winners of round r
    (winners <- outcome[round == r, , drop = FALSE])
    #Find all games winners must hav eown in r-1, ... 1
    rows <- path[winners, 1:(r-1), drop = FALSE]
    # Find which simulaiton to wich each of those games corresponds
    columns <- matrix(rep(1:1, each = nrow(winners)), nrow = length(winners), ncol = r -1 )
    #Replace the corresponding winner with the implied winners
    outcome [cbind(c(rows), c(columns))] <- rep(winners, times = ncol(rows))
}
```



